Chess EDA

### 2023-04-18 owen/Write Ups/01_eda.qmd
some small fixes

Given this output, we can get the probability a player who is black
will win against an equally ranked player, when their current win ratio 
is exactly their average will be $\frac{1}{1+\exp(-\alpha_j)}$.
So a positive $\alpha_j$ indicates a player favoured to win 
even when playing black. The overall $\mu_1$ tells us about the 
global probability of winning as black against an equally ranked
opponent.
If a player
is playing as white their win probability will be
$\frac{1}{1+\exp(-(\alpha_j + \gamma_1))}$,
so $\gamma_1$ being positive indicates the win
probability increases when playing white, which makes sense.
Playing white makes a player who has 50% chance of winning as black
have approximately a 54% chance of winning with white. Wikipedia says the increase 
in win probability from playing white has been estimated to be 
4-6% so quite similar.

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
```

We also consider the previous 20 games.

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1","gamma1", "gamma2")))
```

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))

@@

The size of the winner effects does appear to decrease as you look
at a longer amount of previous games. Unlikely to play more than 
50 games in a row maybe? Still clear positive effects even 
using 50 games.

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))

@@

@@

When we consider more historic games the winner effects do appear to shrink,
as we saw in the bullet data. However, there seems to still be
very strong evidence for some effect. This model does not fit as 
well, but this can be fixed.

@@

- The size of these winner effects decreases as we look at their win ratio over
more historic games.

>>>>

Given this output, we can get the probability a player who is black
will win against an equally ranked player, when their current win ratio 
is exactly their average will be $\frac{1}{1+\exp(-\alpha_j)}$.
So a positive $\alpha_j$ indicates a player favoured to win 
even when playing black. The overall $\mu_2$ tells us about the 
global probability of winning as black against an equally ranked
opponent, which is just below 0, indicating a slightly less than 50% 
chance of winning.
If a player
is playing as white their win probability will be
$\frac{1}{1+\exp(-(\alpha_j + \gamma_1))}$,
so $\gamma_1$ being positive indicates the win
probability increases when playing white, which makes sense.
Playing white makes a player who has 50% chance of winning as black
have approximately a 54% chance of winning with white. Wikipedia says the increase 
in win probability from playing white has been estimated to be 
4-6% so quite similar.

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
```

The posterior of these effects does seem to change somewhat 
as we vary the history. This also seems to vary from player to player.
We also consider the previous 20 games.

@@

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1","gamma1", "gamma2")))

@@

The size of the winner effects does appear to decrease as you look
at a longer amount of previous games. Unlikely to play more than 
50 games in a row maybe? It does appear that the mean
for $\tau_2$ increases as we consider longer histories.

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

@@

When we instead consider blitz games there seems to be considerably less
evidence for these winner effects, with many of the $\beta$ parameters
centered around 0 now. This seems reasonable, indicates
smaller effects in longer games.

@@

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
```

This seems to mirror the smaller number of blitz games quite closely.

@@
